---
title: "Lab 1 Report"
author: "Cory Lanker"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3.25, fig.width = 6.5, fig.align = 'center')
library(ggplot2)
library(latex2exp)
```

## Introduction

This lab explores basic properties of the Si and Ge diodes, including temperature effects on diode current.
SPICE simulations validate lab measurements.

## Diode measurements

The following $I$ vs. $V$ measurements come from
a [1N4148](https://www.digikey.com/en/products/detail/onsemi/1N4148/458603) Si rectifier diode and
for the [1N34](https://www.digikey.com/en/products/detail/nte-electronics-inc/1N34A/11644260) Ge rectifier diode.
SPICE simulations confirmed that the 1N4148 performs similar
to a 1N914 diode, as no SPICE model for the 1N916 could be
located.

```{r p1}
si_iv <- readr::read_csv("p1-si-diode-i-vs-v.csv")
si_iv |>
  ggplot(aes(x = mV, y = mA, color = Sample)) +
  geom_point() +
  geom_line() +
  stat_smooth(method = "lm",
              size = 1/5, alpha = 1/3) +
  labs(x = TeX(r"($V_D$ (mV))"),
       y = TeX(r"($I_D$ (mA))"),
       title = "Diode lab measurements for Si (1N4148) and Ge (1N34A)",
       subtitle = "Linear fit shown for non-linearity referencing") +
  theme(legend.position = 'top') +
  scale_y_log10()
```

The exponential effect of the current is amplified at low
voltages and dampened
at high voltages. The former effect may be due to milliamp-meter low current errors. The latter effect may again
be due to internal diode resistance.

There is something preventing the Schottky exponential behavior
for the Ge diode (1N34A). I don't know why, but posit that it is
due to the current limitations of this diode. Its rated amperage is
50 mA though.

## Comparing with SPICE models

```{r p3}
spice_dat <- readr::read_delim("spice-diodes.txt", delim = " ") |>
  dplyr::select(VD, ISi, IGe) |>
  tidyr::pivot_longer(cols = 2:3, names_to = "Diode")

spice_dat |> 
  ggplot(aes(x = VD*1000, y = value*1000)) +
  geom_line(aes(color = Diode)) +
  geom_point(data = si_iv |>
               dplyr::mutate(Sample = factor(Sample)), 
             mapping = aes(x = mV, y = mA, shape = Sample),
             color = 'grey50',
             size = 1) +
  labs(x = TeX(r"($V_D$ (mV))"),
       y = TeX(r"($I_D$ (mA))"),
       title = "SPICE models for Si diode 1N4148 and Ge Diode 1N34A",
       subtitle = "I-V characteristic")
```

The SPICE model roughly agree with the measurements.

## Temperature effects in SPICE

Here are the I-V curves for different temperatures for these two diodes.
It's reasonable that the current increases with temperature
for a given diode voltage due to the increased 
carrier excitation and mobility.

```{r p5-temp}
dat_temp <- readr::read_table("spice-diodes-temp.txt") |>
  dplyr::rename("V" = "sweep",
                ".1N4148_Si" = "I_Si", 
                ".1N34A_Ge" = "I_Ge") |>
  dplyr::mutate(Temp = rep(c(27, 52, 77), each = 51)) |>
  tidyr::pivot_longer(cols = dplyr::contains("1N"),
                      names_to = "Diode",
                      values_to = "I") |>
  dplyr::select(-Index, -X5)

dat_temp |>
  tidyr::unite(Diode, Temp, col = "Diode_Temp", remove = FALSE) |>
  ggplot(aes(x = V, y = I, color = factor(Temp), linetype = Diode, group = Diode_Temp)) +
  geom_line() +
  labs(x = TeX(r"($V_D$ (V))"),
       y = TeX(r"($I_D$ (A))"),
       title = "I-V characteristics for increasing circuit temperatures in SPICE",
       color = TeX(r"(Temp ($^\circ$C))"))
```

I can get the temperature changes in diode voltage 
for a given current with the following method:

1. for a given diode, calculate the corresponding voltage
for currents $(0.01, 0.02, \dots, I_{max})$ A for
the tested temperatures.
2. plot the changes in voltage for each of the two
temperature increases

Here are plots of 25$^\circ$C changes in voltage.

```{r}
calc_v_given_i <- function(v, i) {
  i_seq <- seq(0.01, max(floor(i * 100))/100, 0.01)
  return(tibble::tibble(i_seq = i_seq,
                        v_seq = stats::approx(i, v, i_seq)$y))
}

dat_del_temp <- dat_temp |>
  tidyr::nest(data = c(V, I)) |>
  dplyr::mutate(y = purrr::map(data, \(x) calc_v_given_i(x$V, x$I)))
```


```{r}
dat_del_temp |>
  dplyr::select(-data) |>
  tidyr::unnest(y) |>
  tidyr::unite(Diode, Temp, col = "Diode_Temp", remove = FALSE) |>
  ggplot(aes(x = i_seq, y = v_seq, color = factor(Temp), linetype = Diode, group = Diode_Temp)) +
  geom_line() +
  labs(y = TeX(r"($V_D$ (V))"),
       x = TeX(r"($I_D$ (A))"),
       title = "V-I characteristics for increasing circuit temperatures in SPICE",
       color = TeX(r"(Temp ($^\circ$C))"))
```

```{r}
# calculate the voltage slope versus temperature within
#  Diode x i_seq
dat_del_temp |>
  dplyr::select(-data) |>
  tidyr::unnest(y) |>
  tidyr::nest(c(Temp, v_seq)) |>
  dplyr::mutate(dV_dT = purrr::map_dbl(
    data, \(x) lm(v_seq ~ Temp, data = x)$coef[["Temp"]]
  )) |>
  ggplot(aes(x = i_seq, y = 1000*dV_dT, color = Diode)) +
  geom_line() +
  labs(y = TeX(r"($\Delta V/\Delta T$ (mV/deg.))"),
       x = TeX(r"($I_D$ (A))"),
       title = "Voltage change per degree for these diodes vs. current",
       subtitle = "Based on SPICE simulations")
```

The voltage sensitivity with temperature seems to increase
with diode current. The slopes are similar between diodes,
about $-1.2$ mV/degree for the 1N34A and $-1.4$ mV/degree for the 1N4148.
