---
title: "Lab 4 Report"
author: "Cory Lanker"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3.25, fig.width = 6.5, fig.align = 'center')
library(ggplot2)
library(latex2exp)
```

<!--
See Figure \ref{fig:p3mod} below.

![Figure caption.](p3mod.pdf){#fig:p3mod width=600 height=300}
-->

## Introduction

This lab introduces the BJT. Tasks in this lab are:

1. 


```{r spice-data}
dat <- readr::read_delim("../ib_1u.txt", delim = " ")
```

The emission coefficient, $n_F$, is a factor included in the
denominator of the diode current vs. voltage equation:

$$ I_D = I_S \exp\left(\frac {V_D} {n_F\, V_T}\right)$$

Some research guided me to typical $n_F$ values, between 1 and 2. 
For IC diodes, $n_F$ is closer to 1, while
with the discrete diodes used in the lab $n_F$ is closer to 2.
I'll show that the emission coefficient is important to make
the small-signal theory align with the measurements.

## Setup

I have a square wave generator producing a DC offset $V_{S}$ and
an AC signal $v_s$. This generator feeds a series combination
of a resistor of measured resistance $98.5 \Omega$ and 8 1N4148
fast-switching diodes. The 1N4148 diode is similar to the requested
1N914 diode. I use eight diodes so I can have a larger AC
signal and still have the small-signal model be valid.
The small-signal model is

$$I_d = I_S \exp\left( \frac{V_D + v_d} {n_F\, V_T} \right)
= I_S \exp\left( \frac{V_D} {n_F\, V_T} \right)
\exp\left( \frac{v_d} {n_F\, V_T} \right)
$$
$$\approx
I_S \exp\left( \frac{V_D} {n_F\, V_T} \right)
\left(1 + \frac{v_d} {n_F\, V_T} \right)$$

where the last approximation is due to the Taylor series expansion
of $exp(x) \approx 1 + x$ when $0 < x << 1$ meaning $v_d$ (the AC signal) magnitude should be much less than 50 mV (since $n_F=2$).

## Measurements

The AC signal measured was the amplitude of the 1kHz square wave about its
DC signal (given in mV in the table). All measurements performed with a
Fluke 175 multimeter. The following table shows the
measurements:

* `V_in_mV`: the DC signal magnitude at node `in`, in mV
* `V_out_mV`: the DC signal magnitude at node `out`, in mV
* `V_in_ac`: the AC signal amplitude (about the DC signal) at node `in`, in mV
* `V_out_ac`: the AC signal amplitude (about the DC signal) at node `out`, in mV
* `I_d_mA`: the current through the eight series diodes, in mA
* `r_pi`: $\frac{n_F\, V_T} {I_D}$, the theoretical $r_\pi$ per diode using $n_F = 2$
* `p_obs`: observed voltage drop proportion for the eight diodes
* `r_est`: observed $r_\pi = \tfrac R 8 \tfrac p {1-p}$ as calculated from the voltage drop proportion $p$
* `obs_err`: percent error in observed $r_\pi$ vs. theoretical values

```{r read-in-data}
R_S <- 98.5
dat <- readr::read_csv("diode-ac-meas.csv") |>
  dplyr::mutate(
    I_d_mA = (V_in_mV - V_out_mV)/R_S,
    r_pi = 2 * 26 / I_d_mA,
    p_obs = V_out_ac/V_in_ac,
    r_est = p_obs / (1-p_obs) * R_S / 8,
    obs_err = (r_est - r_pi)/r_est
  )
column_names <- names(dat)
gt::gt(dat) |> 
  gt::fmt_number(columns = column_names[3:8], n_sigfig = 3) |>
  gt::fmt_percent(columns = "obs_err", decimals = 1)
```

Figure \ref{fig:plot} shows that the error is related to the diode current.
It is interesting that a fit line passes close to zero error when the current
goes to zero.

```{r plot, fig.cap = "Estimated small-signal resistance errors."}
dat |>
  ggplot(aes(x = I_d_mA, y = 100*obs_err)) +
  geom_point() +
  stat_smooth(method = 'lm', fullrange = TRUE) +
  scale_x_continuous(limits = c(0, 7)) +
  labs(x = "Diode current (mA)",
       y = latex2exp::TeX(r"(Observed error in $r_\pi$ (%))"))
```

## Discussion

An emission coefficient of 2 is appropriate for these diodes.
And it seems that the theoretical $r_\pi$ calculation is
valid for small diode currents. With higher diode currents
there appears to be an increase in the diode's small-signal resistance.



